Copulas and their potential for ecology

### 2019-12-18 Paper.Rmd
continuing to trim words here and there

<!--***DAN: got to here with final trim-->

One way we may expect, *a priori*, a study of tail associations and of the 
complex copula dependence structures between
ecological variables
to better reflect ecological relationships relates to Liebig's law of the
minimum. Liebig's law is 
the idea that growth is controlled not by total resources but by the resources which
are scarcest relative to organism needs. 
If, for instance, the growth of a plant depends on soil 
nitrogen, N, and other factors, a plot of plant growth rates versus soil N may 
look more like Fig. \ref{fig_cop_pedag1}A or C than like 
Fig. \ref{fig_cop_pedag1}B or D, i.e., the two variables may show 
left-tail association. Nitrogen 
will control plant growth, producing a clear
relationship, only when it is limiting, in the left portions of the distributions. 
The portions of the distributions that dominate the association are visible in a rank-based
plot, but may not be revealed by standard correlation techniques.

A second reason why we may expect understanding of complex dependence structures
to benefit ecological
research is that prior work has demonstrated, using copulas, complex 
stucture in the spatial dependence of environmental 
variables [@Serinaldi2008; @She2018; @Goswami2018;
@li2013]. One may expect an environmental variable 
measured through time in two locations to show strong tail associations between
the locations if
intense meteorological events are also widespread in their effects, as seems 
frequently to be the case: extreme values are associated with intense events,
so happen in both places at the same time, whereas moderate values of the 
environmental variable may be associated with local phenomena, which differ
between the locations.
Spatial dependence in an environmental variable 
tends to beget spatial dependence 
in fluctuations of populations or other ecosystem variables 
(this is called \emph{spatial synchrony}) which are influenced by the environmental 
variable. This influence is called the \emph{Moran effect}. 
If asymmetric tail associations, or other complex dependence structure, is transmitted, 
possibly in modified form, 
from environmental to ecological variables, then we may expect complex dependence 
structure and tail associations to be a common feature of the 
spatial synchrony of population, community, 
biogeochemical and other environmentally influenced ecological variables.
Synchrony is a phenomenon of substantial interest in 
ecology [@Liebhold2004; @Sheppard2016; @walteretal2017].

A third reason why we may expect an understanding of complex 
dependence structures and tail associations to benefit 
ecological research is that such an approach may help illuminate 
the causal mechanisms between variables. For instance, if two species
Sp1 and Sp2 are strong competitors, abundances of the two species
across quadrats should be negatively related, as in 
Fig. \ref{fig_cop_pedag1}E-F. If Sp1 is plotted on the horizontal axis
and Sp2 on the vertical, then Fig. \ref{fig_cop_pedag1}E 
may suggest Sp1 is the dominant 
competitor: when Sp1 is abundant, Sp2 is necessarily rare because it is suppressed; 
whereas when Sp1 is rare, Sp2
may be abundant, or may also be relatively rare due to 
limiting factors other than Sp1. 
Alternatively, Fig. \ref{fig_cop_pedag1}F may suggest that Sp2
is the dominant competitor. Other causal hypotheses may produce 
similar dependence structure, and it is usually impossible to obtain
complete information on causal pathways from analyses which are
fundamentally based on associations. However, more detailed 
examination of the tail associations and potentially other 
patterns of dependence of variables may
well be combined with biological and other reasoning to rule out
some causal hypotheses which could not be eliminated solely 
with standard correlation approaches. @Popovic2018 has previously
used copulas to help illuminate causal aspects of relationships
between species.

A fourth reason why we may expect an understanding of tail associations and of complex 
dependence structures to be useful
for ecology has to do with spatially aggregated or averaged quantities. 
Many ecological variables of applied importance 
depend on the spatial average or sum of local 
quantities. For instance, regional methane and CO$_2$ 
fluxes are the sum of local fluxes; 
and the total economic value of a fishery is the sum of local catches. 
We will explore how details, which can be captured using copulas and which
relate to tail associations, 
of the dependence between fluctuations of local
quantities can influence fluctuations of 
the spatial mean or sum, and how this may influence 
higher organizational levels in ecology and human concerns.
To illustrate this idea we cite @li2013, who have demonstrated that the overall 
reliability of wind-generated electricity depends sensitively on details
of the dependence between wind speeds at multiple generator sites. 
Spatially aggregated ecological variables may be subject to a similar 
effect, with
possible consequences. For instance, if populations of a 
pest species in different locations are all positively associated and are also more strongly
related to each other in their right tails, then local outbreaks will tend 
to occur together, creating regional epidemics. 
On the other hand, stronger left-tail associations in a pest species, even if overall 
correlation were unchanged, would have more benign effects.

We approached our overall goal of exploring the potential of copula approaches 
for ecology by addressing the following specific questions. (Q1) Do datasets
in ecology have dependence structure distinct from that of a multivariate 
Gaussian/normal
distribution (here called non-normal copula structure)? Do positively associated 
ecological variables show tail 
associations distinct from those of a normal distribution, and in particular
do they show asymmetric tail associations? Normal copula structure is assumed
by standard approaches that use multivariate 
normal distributions or distributions obtained by transforming
the marginals of a normal distribution. So Q1 asks whether ecological
data contain dependency information distinct from a standard or default case. 
(Q2) If the answer to
Q1 is "yes," then what are some possible causes/mechanisms of non-normal
copula structure and asymmetric tail associations in ecology? And (Q3), what are the 
consequences of non-normal copula
structure and asymmetric tail associations for ecological understanding and
applications? After section \ref{Background}, 'Background on copulas', we address Q1 
(sections \ref{Data} through \ref{Results} of the paper) 
by analyzing several datasets, including 
environmental, species-trait, phenological, population, community, and 
ecosystem functioning data, selected to span multiple 
sub-fields and organizational levels of ecology. 
We address Q2 (sections \ref{Causes} through \ref{Results_Q2} of the paper), using 
simple models. We address Q3 (sections \ref{Consequences} through \ref{Results_Q3} 
of the paper) using 
both data and models. Section \ref{Discussion} is the Discussion.
Multiple analyses are brought to bear on each question, and these
are summarized diagramatically in Fig. \ref{MasterFigure}, which can 
serve as a *post hoc* guide to the paper. 
Copula approaches have been used to great effect in neuroscience [@onken2009], 
bioinformatics [@Kim2008], medical research [@Emura2016], direct study of environmental 
variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], and finance [@Li2000],
and have also been used very effectively, but rarely so far, in ecology [@Valpine2014; 
@Anderson2018; @Popovic2018]. 
We argue that benefits of wider usage of complete 
copula descriptions of dependence in ecology will be substantial.

@@

Sklar's theorem implies that
any random vector can be constructed from a unique copula and marginal 
cdfs, and, furthermore,
any copula and any univariate cdfs give rise to a random vector;
so Sklar's theorem provides a correspondence between random vectors
and pairs consisting of a copula and two univariate cdfs,
thereby making it possible to study a random vector by studying these two
constituents.
Marginals contain no information about the
dependence structure of a random vector, so the copula contains
all such information - it is a complete description of the dependence
between variables.
<!--***-->
The univariate cdfs associated with a random vector 
$(X,Y)$ are its margins, $F_X$ and $F_Y$, and the 
associated copula is the copula, $C$, guaranteed by Sklar's theorem
(and guaranteed unique, thanks to the assumptions of continuity 
and monotonicity made above). In fact, and making Sklar's correspondence
more concrete, we show in Appendix \ref{SM-ContStrictMonoton}
that $C$ is the bivariate distribution
function of the random variable $(F_X(X),F_Y(Y))$. 
<!--***-->
Conversely, given a copula
$D$ and univariate cdfs $G_A$ and $G_B$,
Sklar's theorem guarantees that $D(G_A(x),G_B(y))$ is a 
cdf, so the random variable with this 
cdf is the random variable corresponding
to $D$, $G_A$ and $G_B$ under Sklar's correspondence. 
Again making Sklar's correspondence
more concrete, we show in Appendix \ref{SM-ContStrictMonoton}
that this random variable is $(G_A^{-1}(U),G_B^{-1}(V))$, 
where $G_A^{-1}$ and $G_B^{-1}$ are the inverses of $G_A$ and $G_B$
and $(U,V)$ is the random variable with cdf $D$
(which has uniform marginals).
If we conflate a copula with the random vector of 
which the copula is the cdf, then Sklar's correspondence
is between random variables, $(U,V)$, with uniform marginals 
and random variables, $(X,Y)$, with arbitrary continuous, strictly monotonic
marginals. The correspondence is simply via the application of the
univariate cdfs, or their inverses: $U=F_X(X)$,
$V=F_Y(Y)$, $X=F^{-1}_X(U)$, $Y=F^{-1}_Y(V)$.
Application of the 
cdfs of the component random variables translates
between the uniform-marginal context, $(U,V)$, which isolates the 
dependence information, and the context of the original random variable,
$(X,Y)$.

Thus Sklar's theorem makes it possible to construct
bivariate distributions in two separate steps: by specifying the marginal 
distributions and by specifying the dependence structure. For instance,
construction of several bivariate distributions is carried out in this way in
Fig. \ref{PedagogFigCopThry}. The contrast between Fig. \ref{PedagogFigCopThry}C and D
illustrates in particular how a bivariate distribution can be changed, while retaining 
the same marginals, by changing the copula. Fig. \ref{PedagogFigCopThry}C is a 
bivariate normal distribution, but Fig. \ref{PedagogFigCopThry}D clearly 
is not, although its marginals
are the same. Fig. \ref{PedagogFigCopThry}D shows stronger association of the two variables
in the left than in the right tails. Sklar's theorem
also makes it possible to study and model dependence separately 
from marginal distributions. For instance, in a pedagogical 
example, @Anderson2018 modelled the dependence between abundance
data for two fish species, red moki and black angelfish, by first 
modelling the marginal distributions and then modelling 
the dependence using Gaussian copulas (their section 2).
See the Discussion for a few words on multivariate copulas, which can
be developed in the same way as above [@nelsen2006_copula; @joe2014_dependence;
@MaiScherer2017; @Anderson2018].

@@

We introduce some standard copula families, as examples of copulas and also
because we will fit these families to data later. The bivariate normal
copula family, which consists of copulas of bivariate normal
distributions, is a one-parameter family. The parameter, which we here call $p$,
corresponds to the correlation
of the related bivariate normal distribution, and controls the degree of association
between the variables. A normal copula with $p=0.7$ was already introduced (Fig.
\ref{PedagogFigCopThry}A). The formula for the
copula is $\Phi_{2,p}(\Phi_{1}^{-1}(u),\Phi_{1}^{-1}(v))$, where $\Phi_1^{-1}$ is the inverse of the
cummulative dstribution function (cdf) of a (univariate) standard 
normal and $\Phi_{2,p}$ is the cdf of a
bivariate normal distribution with mean $(0,0)$ and covariance matrix having $1$s on the diagonal
and $p$ in the off diagonal positions. Formulas for the normal copula and all the 
other copula families we use were provided by @Brechmann2013. 
However, formulas for copulas are often not as 
directly informative about copula properties as probability density function (pdf) 
pictures; Fig. \ref{figped_N} has 
example pdfs corresponding to bivariate normal copulas for various values of $p$.
Note that the pdfs are symmetric
around the diagonal line $v=-u+1$ (Fig. \ref{figped_N}), so normal copulas have symmetric 
associations between the two variables in the left and right tails. 
The Clayton copula family,
of which we have already pictured an example (Fig.
\ref{PedagogFigCopThry}B), is another one-parameter family. In contrast to normal copulas, Clayton copulas
have stronger left- than right-tail association. The formula is 
$\left[  \reumax(u^{-p}+v^{-p}-1,0)\right]^{-1/p}$, for parameter $p$; though this again provides 
less direct intuition than do example pdfs (Fig. \ref{figped_C}). Higher values of the parameter,
$p$ produce copulas from the Clayton family with higher Kendall or Spearman correlation. The survival
Clayton family is the symmetric opposite of the Clayton family, showing
stronger right- than left tail association (Fig. \ref{SM-figped_SC}). The BB1 copula 
family is a two-parameter family, thereby providing more flexibility: for some parameters
BB1 copulas have stronger left- than right-tail association, and for others the reverse
(Fig. \ref{figped_BB1}). See @nelsen2006_copula, @joe2014_dependence and 
@Brechmann2013 for further
information on these and other copula families.

Having introduced tail association conceptually and referred to it
in examples, we now need a more precise definition of a measure of tail association.
We use the term *tail association* to describe the general idea of the strength of
associations between two variables in the tails of their distributions. One of the
ways this is measured is called *tail dependence* (@joe2014_dependence, section 2.13; 
@nelsen2006_copula, section 5.4), which we here define. Given a random vector $(X,Y)$
with margins $F_X$, $F_Y$ and defining $U=F_X(X)$, $V=F_Y(Y)$, the upper-tail
dependence of $X$ and $Y$ is defined as 
$\lambda_U = \lim_{u \rightarrow 1^-} P[Y>F_Y^{-1}(u) | X>F_X^{-1}(u)]$. This equals
$\lim_{u \rightarrow 1^-} P[U>u | V>u]$, which in turn equals
$\lim_{u \rightarrow 1^-} P[U>u, V>u]/P[V>u]=\lim_{u \rightarrow 1^-} P[U>u, V>u]/(1-u)$,
which shows upper-tail dependence is defined symmetrically in the two random variables.
All the variables we consider were positively associated when they were significantly
associated, so we think of positively associated variables when conceptualizing the
definitions presented here. Lower tail dependence is defined analogously, as
$\lambda_L = \lim_{u \rightarrow 0^+} P[Y \leq F_Y^{-1}(u) | X \leq F_X^{-1}(u)]$.
Tail dependence is a property of the random variable $(U,V)$, so depends only
on the copula and not on the marginals combined with it. It can be shown 
(@nelsen2006_copula, section 5.4) that $\lambda_L=\lambda_U=0$ for the normal copula,
$\lambda_L=2^{-1/p}$ and $\lambda_U=0$ for a Clayton copula with parameter $p>0$,
and $\lambda_L=0$ and $\lambda_U=2^{-1/p}$ for a survival Clayton  copula with parameter $p>0$.
In section \ref{Methods} we will introduce
ways tail dependence can be applied to data, by first fitting copulas to the data. We will also 
introduce nonparametric measures of tail association that apply directly to data.

Plots using normalized ranks were proposed in the Introduction as being conceptually
similar to copulas. We here explain the connection. As stated above, the copula associated
with a random variable $(X,Y)$ is the cdf of the random variable
$(F_X(X),F_Y(Y))$. Given a sample $(x_i,y_i)$, $i=1,\ldots,n$, from $(X,Y)$, let $\hat{F}_X$ and
$\hat{F}_Y$ be the empirical cdfs associated with the $x_i$ and $y_i$, respectively.
These are step functions of $x$ and $y$, respectively, that start at $0$ for low values of 
$x$ and $y$ and jump by $1/n$
at each of the data points. Therefore $\hat{F}_X(x_j)$ is the rank of $x_j$ in
the set $\{x_i : i=1,\ldots,n\}$, dividided by $n$, 
and analogously $\hat{F}_Y(y_j)$ is the rank of $y_j$ in the set 
$\{y_i : i=1,\dots,n\}$, divided
by $n$. For large $n$, $\hat{F}_X(x_i)$ and $\hat{F}_Y(y_i)$ 
approximate the cdfs $F_X$ and $F_Y$, respectively. 
But $\hat{F}_X(x_i)$ equals the normalized rank of $x_i$ times
$n/(n+1)$, and likewise $\hat{F}_Y(y_i)$ equals the normalized rank of $y_i$ times
$n/(n+1)$. So the normalized rank in turn approximates the empirical cdf and therefore 
the cdf. Thus the normalized rank pairs $(u_i,v_i)$ for $i=1,\ldots,n$ can be 
regarded as approximate samples from the random variable $(F_X(X),F_Y(Y))$, 
which is the random variable associated with the copula of $(X,Y)$.
The scatterplot of $v_i$ versus $u_i$ can be used to infer aspects of copula 
structure (see section \ref{Methods} below).

@@

Recall that Q1 has three parts: Do datasets
in ecology have dependence structure distinct from that of a multivariate 
Gaussian/normal
distribution (here called non-normal copula structure)? Do positively associated 
ecological variables show tail 
associations distinct from those of a normal distribution? And in particular
do they show asymmetric tail associations? 
We addressed Q1 first via a model selection procedure in which several
families of copulas were fitted to 
our ecological datasets and fits were compared via the Akaike and Bayesian 
Information Criteria (AIC and BIC). One of the fitted copulas was the normal
copula, making possible comparisons of the degree to which the normal copula versus
other copula families were good descriptions of data.

For bivariate datasets 
$(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps. 
First, we produced normalized ranks 
$u_i$ and $v_i$ as in the Introduction. 
Second, we tested the independence of the $u_i$ and $v_i$
using the statistic $T=\sqrt{(9N(N-1))/(4N+10)}|\tau|$, where $\tau$
is Kendall's tau for the data. @Genest2007 argue that this statistic is approximately standard normally
distributed. We used the 
implementation of this test given in the `BiCopIndTest` in the `VineCopula` package
in the R programming language. 
We tested for independence because our model selection 
algorithms were ineffective if data could not 
be distinguished from independent data 
since many of the copula families we considered include the independent copula. 
If independence could be rejected ($0.05$ significance level), model 
selection proceeded. Third, we fit 16 bivariate copula families (see below for 
information on the families used) to the 
normalized ranks via maximum likelihood. The approach of fitting
copula families to normalized ranks was studied in depth
and recommended by @Genest1995 and @Shih1995. Their estimator of copula
parameters, which we use, was shown to be consistent, asymptotically normal,
and fully efficient at indepencence [@Genest1995]. @Genest2007
recommend carrying out inferences of dependence structures (which was our
goal here) using normalized ranks.
We used the fitting implementation given in `BiCopEst` in `VineCopula`.
Fourth, we obtained AIC and BIC values and accompanying model weights 
$\text{AIC}_{\text{w}}$ and 
$\text{BIC}_{\text{w}}$ [@burnham2003_modelselection] for each fitted copula. 
`BiCopEst` also 
provided lower- and 
upper-tail dependence of the best-fitting member of each family. 
$\text{AIC}_{\text{w}}$ values were used to get
model-averaged lower- and upper-tail dependence values using standard model averaging 
formulas [@burnham2003_modelselection]; likewise for BIC.

>>>>

One way we may expect, *a priori*, a study of tail associations and 
copula dependence structures 
to better reflect ecological relationships relates to Liebig's law of the
minimum. Liebig's law is 
the idea that growth is controlled not by total resources but by the resource which
is scarcest relative to organism needs. 
If, for instance, the growth of a plant depends on soil 
nitrogen, N, and other factors, a plot of growth rates versus soil N may 
look more like Fig. \ref{fig_cop_pedag1}A or C than like 
Fig. \ref{fig_cop_pedag1}B or D, i.e., the two variables may show 
left-tail association: N 
controls plant growth, producing a clear
relationship, only when it is limiting. 
This aspect of the association would be visible in a rank-based
plot, but may not be revealed by correlation.

A second reason why we may expect detailed understanding of dependence structures
to benefit ecological
research is that prior work demonstrated, using copulas, complex 
stucture in the spatial dependence of environmental 
variables [@Serinaldi2008; @She2018; @Goswami2018;
@li2013]. An environmental variable 
measured through time in two locations may show strong tail associations between
the locations if
intense meteorological events are also widespread, as seems 
frequently to be the case: extreme values are associated with intense events,
so happen in both places at the same time, whereas moderate values of the 
environmental variable may be associated with local phenomena, which differ
between the locations.
Spatial dependence in an environmental variable 
tends to beget spatial dependence 
in fluctuations of populations or other ecosystem variables 
(this is called \emph{spatial synchrony}) influenced by the environmental 
variable. This is called the \emph{Moran effect}. 
If asymmetric tail associations, or other complex dependence structure, is transmitted 
from environmental to ecological variables, then we would expect complex dependence 
structure and tail associations to be a common feature of the 
spatial synchrony of population, community, 
biogeochemical and other environmentally influenced ecological variables.
Synchrony attracts major interest in 
ecology [@Liebhold2004; @Sheppard2016; @walteretal2017].

A third reason why we may expect an understanding of complex 
dependence structures and tail associations to benefit 
ecology is that such an approach may help illuminate 
causal mechanisms between variables. If two species,
Sp1 and Sp2, are strong competitors, abundances of the two species
across quadrats should be negatively related, as in 
Fig. \ref{fig_cop_pedag1}E-F. If Sp1 is plotted on the horizontal axis
and Sp2 on the vertical, then Fig. \ref{fig_cop_pedag1}E 
may suggest Sp1 is the dominant 
competitor: when Sp1 is abundant, Sp2 is necessarily rare because it is suppressed; 
whereas when Sp1 is rare, Sp2
may be abundant, or may also be relatively rare due to 
limiting factors other than Sp1. 
Alternatively, Fig. \ref{fig_cop_pedag1}F may suggest Sp2
is the dominant competitor. Other causal hypotheses may produce 
similar dependence structure, and it is usually impossible to obtain
complete information on causal pathways from analyses which are
fundamentally based on associations. However, 
examination of tail associations may
be combined with biological reasoning to rule out
some causal hypotheses which could not previously be eliminated. 
@Popovic2018 used copulas to illuminate causal relationships
between species.

A fourth reason why we may expect an understanding of tail associations and complex 
dependence structures to be useful
for ecology has to do with spatially aggregated or averaged quantities. 
Many ecological variables of applied importance 
depend on the spatial average or sum of local 
quantities. For instance, regional methane and CO$_2$ 
fluxes are the sum of local fluxes; 
and the total economic value of a fishery is the sum of local catches. 
We will explore how tail associations between fluctuations of local
quantities can influence fluctuations of 
the spatial mean or sum, and how this may influence 
higher organizational levels in ecology and human concerns.
To illustrate this idea we cite @li2013, who demonstrated that the overall 
reliability of wind-generated electricity depends sensitively on details
of the dependence between wind speeds at multiple generator sites. 
Spatially aggregated ecological variables may be subject to a similar 
effect. For instance, if populations of a 
pest species in different locations are all positively associated and are also more strongly
related to each other in their right tails, then local outbreaks will tend 
to occur together, creating regional epidemics. 
Stronger left-tail associations in a pest species, even if overall 
correlation were the same, would have more benign effects.

We approached our overall goal of exploring the potential of copula approaches 
for ecology by addressing the following specific questions. (Q1) Do datasets
in ecology have dependence structure distinct from that of a multivariate 
Gaussian/normal
distribution (here called non-normal copula structure)? Do positively associated 
ecological variables show tail 
associations distinct from those of a normal distribution, and in particular
do they show asymmetric tail associations? Normal copula structure is assumed
by standard approaches that use multivariate 
normal distributions or distributions obtained by transforming
the marginals of a normal distribution. So Q1 asks whether ecological
data contain dependency information distinct from a standard or default case. 
(Q2) If the answer to
Q1 is "yes," then what are some possible causes/mechanisms of non-normal
copula structure and asymmetric tail associations in ecology? And (Q3), what are the 
consequences of non-normal copula
structure and asymmetric tail associations for ecological understanding and
applications? After section \ref{Background}, 'Background on copulas', we address Q1 
(sections \ref{Data} through \ref{Results} of the paper) 
by analyzing several datasets, including 
environmental, species-trait, phenological, population, community, and 
ecosystem functioning data, selected to span multiple 
sub-fields and organizational levels of ecology. 
We address Q2 (sections \ref{Causes} through \ref{Results_Q2}), using 
simple models. We address Q3 (sections \ref{Consequences} through \ref{Results_Q3}) using 
both data and models. Section \ref{Discussion} is the Discussion.
Multiple analyses are brought to bear on each question, and these
are summarized diagramatically in Fig. \ref{MasterFigure}, which can 
serve as a *post hoc* guide to the paper. 
Copula approaches have been used to great effect in neuroscience [@onken2009], 
bioinformatics [@Kim2008], medical research [@Emura2016], direct study of environmental 
variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], and finance [@Li2000],
and have also been used effectively, but rarely so far, in ecology [@Valpine2014; 
@Anderson2018; @Popovic2018]. 
We argue that benefits of wider use of
copula descriptions of dependence in ecology will be substantial.

@@

Sklar's theorem implies that
any random vector can be constructed from a unique copula and marginal 
cdfs, and, furthermore,
any copula and any univariate cdfs give rise to a random vector;
so Sklar's theorem provides a correspondence between random vectors
and pairs consisting of a copula and two univariate cdfs,
thereby making it possible to study a random vector by studying these two
constituents.
Marginals contain no information about the
dependence structure of a random vector, so the copula contains
all such information - it is a complete description of the dependence
between variables.
<!--***-->
The univariate cdfs associated with a random vector 
$(X,Y)$ are its margins, $F_X$ and $F_Y$, and the 
associated copula is the copula, $C$, guaranteed by Sklar's theorem
(and guaranteed unique, thanks to the assumptions of continuity 
and monotonicity made above). In fact, and making Sklar's correspondence
more concrete, we show in Appendix \ref{SM-ContStrictMonoton}
that $C$ is the bivariate distribution
function of the random variable $(F_X(X),F_Y(Y))$. 
<!--***-->
Conversely, given a copula
$D$ and univariate cdfs $G_A$ and $G_B$,
Sklar's theorem guarantees that $D(G_A(x),G_B(y))$ is a 
cdf, so the random variable with this 
cdf is the random variable corresponding
to $D$, $G_A$ and $G_B$ under Sklar's correspondence. 
Again making Sklar's correspondence
concrete, we show in Appendix \ref{SM-ContStrictMonoton}
that this random variable is $(G_A^{-1}(U),G_B^{-1}(V))$, 
where $G_A^{-1}$ and $G_B^{-1}$ are the inverses of $G_A$ and $G_B$
and $(U,V)$ is the random variable with cdf $D$
(which has uniform marginals).
If we conflate a copula with the random vector of 
which the copula is the cdf, then Sklar's correspondence
is between random variables, $(U,V)$, with uniform marginals 
and random variables, $(X,Y)$, with arbitrary continuous, strictly monotonic
marginals. The correspondence is simply via the application of the
univariate cdfs, or their inverses: $U=F_X(X)$,
$V=F_Y(Y)$, $X=F^{-1}_X(U)$, $Y=F^{-1}_Y(V)$.

Thus Sklar's theorem makes it possible to construct
bivariate distributions in two separate steps: by specifying the marginal 
distributions and by specifying the dependence structure.
Construction of several bivariate distributions is carried out in this way in
Fig. \ref{PedagogFigCopThry}. The contrast between Fig. \ref{PedagogFigCopThry}C and D
illustrates in particular how a bivariate distribution can be changed, while retaining 
the same marginals, by changing the copula. Fig. \ref{PedagogFigCopThry}C is a 
bivariate normal distribution, but Fig. \ref{PedagogFigCopThry}D clearly 
is not, although its marginals
are the same. Fig. \ref{PedagogFigCopThry}D shows stronger association of the two variables
in the left than in the right tails. Sklar's theorem
also makes it possible to study and model dependence separately 
from marginal distributions. For instance, in a pedagogical 
example, @Anderson2018 modelled the dependence between abundance
data for two fish species, red moki and black angelfish, by first 
modelling the marginal distributions and then modelling 
the dependence using Gaussian copulas (their section 2).
See the Discussion for a few words on multivariate copulas, which can
be developed in the same way as above [@nelsen2006_copula; @joe2014_dependence;
@MaiScherer2017; @Anderson2018].

@@

We introduce some standard copula families, as examples, and also
because we will fit these families to data. The bivariate normal
copula family, which consists of copulas of bivariate normal
distributions, is a one-parameter family. The parameter, $p$,
corresponds to the correlation
of the related bivariate normal distribution, and controls the degree of association
between the variables. A normal copula with $p=0.7$ was already introduced (Fig.
\ref{PedagogFigCopThry}A). The formula for the
copula is $\Phi_{2,p}(\Phi_{1}^{-1}(u),\Phi_{1}^{-1}(v))$, where $\Phi_1^{-1}$ is the inverse of the
cdf of a (univariate) standard 
normal and $\Phi_{2,p}$ is the cdf of a
bivariate normal distribution with mean $(0,0)$ and covariance matrix having $1$s on the diagonal
and $p$ in the off diagonal positions. Formulas for all the 
copula families we use were provided by @Brechmann2013. 
However, formulas for copulas are often not as 
directly informative about copula properties as probability density function (pdf) 
pictures; Fig. \ref{figped_N} has 
example pdfs corresponding to bivariate normal copulas for various values of $p$.
Note that the pdfs are symmetric
around the diagonal line $v=-u+1$ (Fig. \ref{figped_N}), so normal copulas have symmetric 
associations between the two variables in the left and right tails. 
The Clayton copula family,
of which we already pictured an example (Fig.
\ref{PedagogFigCopThry}B), is another one-parameter family. In contrast to normal copulas, Clayton copulas
have stronger left- than right-tail association. The formula is 
$\left[  \reumax(u^{-p}+v^{-p}-1,0)\right]^{-1/p}$, for parameter $p$; though this again provides 
less direct intuition than do example pdfs (Fig. \ref{figped_C}). Higher values of the parameter,
$p$ produce copulas with higher Kendall or Spearman correlation. The survival
Clayton family is the symmetric opposite of the Clayton family, showing
stronger right- than left tail association (Fig. \ref{SM-figped_SC}). The BB1 copula 
family is a two-parameter family, thereby providing more flexibility: for some parameters
BB1 copulas have stronger left- than right-tail association, and for others the reverse
(Fig. \ref{figped_BB1}). See @nelsen2006_copula, @joe2014_dependence and 
@Brechmann2013 for further
information on these and other families.

Having introduced tail association conceptually and referred to it
in examples, we now need a precise definition of a measure of tail association.
We use the term *tail association* to describe the general idea of the strength of
association between two variables in the tails of their distributions. One 
way this is measured is called *tail dependence* (@joe2014_dependence, section 2.13; 
@nelsen2006_copula, section 5.4), defined here. Given a random vector $(X,Y)$
with margins $F_X$, $F_Y$ and defining $U=F_X(X)$, $V=F_Y(Y)$, the upper-tail
dependence of $X$ and $Y$ is defined as 
$\lambda_U = \lim_{u \rightarrow 1^-} P[Y>F_Y^{-1}(u) | X>F_X^{-1}(u)]$. This equals
$\lim_{u \rightarrow 1^-} P[U>u | V>u]$, which in turn equals
$\lim_{u \rightarrow 1^-} P[U>u, V>u]/P[V>u]=\lim_{u \rightarrow 1^-} P[U>u, V>u]/(1-u)$,
which shows upper-tail dependence is defined symmetrically in the two variables.
All the variables we consider were positively associated when they were significantly
associated, so we think of positively associated variables when conceptualizing the
definitions here. Lower tail dependence is defined analogously, as
$\lambda_L = \lim_{u \rightarrow 0^+} P[Y \leq F_Y^{-1}(u) | X \leq F_X^{-1}(u)]$.
Tail dependence is a property of $(U,V)$, so depends only
on the copula and not on the marginals combined with it. It can be shown 
(@nelsen2006_copula, section 5.4) that $\lambda_L=\lambda_U=0$ for the normal copula,
$\lambda_L=2^{-1/p}$ and $\lambda_U=0$ for a Clayton copula with parameter $p>0$,
and $\lambda_L=0$ and $\lambda_U=2^{-1/p}$ for a survival Clayton  copula with parameter $p>0$.
In section \ref{Methods} we will introduce
ways tail dependence can be applied to data, by first fitting copulas to the data. We will also 
introduce nonparametric measures of tail association.

Plots using normalized ranks were proposed in the Introduction as being conceptually
similar to copulas; we explain the connection. As stated above, the copula associated
with a random variable $(X,Y)$ is the cdf of the random variable
$(F_X(X),F_Y(Y))$. Given a sample $(x_i,y_i)$, $i=1,\ldots,n$, from $(X,Y)$, let $\hat{F}_X$ and
$\hat{F}_Y$ be the empirical cdfs associated with the $x_i$ and $y_i$, respectively.
These are step functions of $x$ and $y$, respectively, that start at $0$ for low values of 
$x$ and $y$ and jump by $1/n$
at each of the data points. Therefore $\hat{F}_X(x_j)$ is the rank of $x_j$ in
the set $\{x_i : i=1,\ldots,n\}$, dividided by $n$, 
and analogously $\hat{F}_Y(y_j)$ is the rank of $y_j$ in the set 
$\{y_i : i=1,\dots,n\}$, divided
by $n$. For large $n$, $\hat{F}_X(x_i)$ and $\hat{F}_Y(y_i)$ 
approximate the cdfs $F_X$ and $F_Y$, respectively. 
But $\hat{F}_X(x_i)$ equals the normalized rank of $x_i$ times
$n/(n+1)$, and likewise $\hat{F}_Y(y_i)$ equals the normalized rank of $y_i$ times
$n/(n+1)$. So the normalized rank in turn approximates the empirical cdf and therefore 
the cdf. Thus the normalized rank pairs $(u_i,v_i)$ for $i=1,\ldots,n$ can be 
regarded as approximate samples from the random variable $(F_X(X),F_Y(Y))$, 
which is the random variable associated with the copula of $(X,Y)$.
The scatterplot of $v_i$ versus $u_i$ can be used to infer aspects of copula 
structure (see section \ref{Methods} below).

@@

Recall that Q1 has three parts: Do datasets
in ecology have non-normal copula structure? Do positively associated 
ecological variables show tail 
associations distinct from those of a normal distribution? And in particular
do they show asymmetric tail associations? 
We addressed Q1 first via a model selection procedure in which several
families of copulas were fitted to 
our ecological datasets and fits were compared via the Akaike and Bayesian 
Information Criteria (AIC and BIC). One of the fitted copulas was the normal
copula, making possible comparisons of the degree to which the normal versus
other copula families were good descriptions of data.

For bivariate datasets 
$(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps. 
First, we produced normalized ranks 
$u_i$ and $v_i$ as in the Introduction. 
Second, we tested the independence of the $u_i$ and $v_i$
using the statistic $\sqrt{(9N(N-1))/(4N+10)}|\tau|$, where $\tau$
is Kendall's tau for the data. @Genest2007 argue that this statistic is approximately standard normally
distributed. We used the 
implementation of this test in `BiCopIndTest` in the `VineCopula` package
in R. We tested for independence because our model selection 
algorithms were ineffective if data could not 
be distinguished from independent data, 
since many copula families include the independent copula. 
If independence could be rejected ($0.05$ level), model 
selection proceeded. Third, we fit 16 bivariate copula families (see below) to the 
normalized ranks via maximum likelihood. The approach of fitting
copula families to normalized ranks was recommended 
by @Genest1995 and @Shih1995. Their estimator of copula
parameters, which we use, is consistent, asymptotically normal,
and fully efficient at indepencence [@Genest1995]. @Genest2007
recommend carrying out inferences of dependence structures (which was our
goal here) using normalized ranks.
We used the fitting implementation given in `BiCopEst` in `VineCopula`.
Fourth, we obtained AIC and BIC values and accompanying model weights 
$\text{AIC}_{\text{w}}$ and 
$\text{BIC}_{\text{w}}$ [@burnham2003_modelselection] for each fitted copula. 
`BiCopEst` also 
provided lower- and 
upper-tail dependence of the best-fitting member of each family. 
$\text{AIC}_{\text{w}}$ values were used to get
model-averaged lower- and upper-tail dependence values using standard model averaging 
formulas [@burnham2003_modelselection]; likewise for BIC.

Thus for each bivariate dataset, 
the end products of our procedure were threefold, corresponding to the
three parts of Q1 listed in the Introduction: A) the AIC, BIC, $\text{AIC}_{\text{w}}$ and 
$\text{BIC}_{\text{w}}$ values for each of our 16 copula families (see below for list),
including the normal family, providing an inference as to whether the normal
copula or an alternative was better supported by 
data; B) lower- and upper-tail dependence measures for each fitted copula and 
model-averaged tail dependence measures, providing information on whether,
and to what extent, tail dependence differed from the tail dependence of 
a normal copula (i.e., 0); and C) the difference of lower- and upper-tail
dependence measures for each fitted copula and model averages of those 
quantities, providing information on whether tail
dependence was asymmetric.
